---
title: "Data Analysis"
author: "AG35441 - Kamila Kwiatkowska (k25120717)"
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
  )

options (scipen = 999)

library(tidyverse)
library(here)        
library(janitor)     
library(psych)       
library(rstatix)     
library(effectsize)  
library(car)         
library(lmtest)      
library(broom)
library(sandwich)  
library(ordinal)   
library(brant) 

```

```{r read-in-data}

survey_data <- read.csv("data/Analysis.csv")

dat <- survey_data |> 
  
clean_names() |>
  rename(
    awareness             = total_awareness_score,
    behaviour             = average_behaviour,
    perceived_control     = s3q1,
    value_recognition     = s3q2,
    platform_transparency = s3q3,
    trust                 = trust_score,
    regulation            = average_regulation_score,
    age                   = age_recoded,
    gender                = gender_recoded,
    education             = education_recoded,
    screen_time           = screen_time_recoded
  ) |>

mutate(
    age_band    = factor(age, levels = 1:5,
                         labels = c("18-24", "25-34", "35-44", "45-54", "55+"),
                         ordered = TRUE),
    
    edu_band    = factor(education, levels = 1:3,
                         labels = c("Secondary", "Undergraduate", "Postgraduate"),
                         ordered = TRUE),
    
    screen_band = factor(screen_time, levels = 1:4,
                         labels = c("<1h", "1-3h", "4-5h", ">5h"),
                         ordered = TRUE),
    
    gender_f    = factor(gender, levels = c(0, 1), 
                         labels = c("Male", "Female"))
    
  ) 

glimpse(dat)
cat("\nN =", nrow(dat), "\n")
    
```

```{r read-in-check}
dat |> tabyl(age_band)
dat |> tabyl(edu_band)
dat |> tabyl(screen_band)
dat |> tabyl(gender_f)
```
#Survey
This is a test analysis of a dataset on Everyday Users' Perceptions of Data Use and Value Creation in the Hidden Value Chains of the Data Economy.

#Hypotheses

##Hypothesis 1
H1 Greater awareness of data extraction is not associated with greater privacy-protective behaviour.

##Hypothesis 2
H2 Greater awareness of data extraction is associated with lower platform trust.

##Hypothesis 3
H3 Greater awareness of data extraction is associated with stronger support for regulation.

##Hypothesis 4
H4 Users recognise their everyday online activity as a source of economic value for platforms.

##Hypothesis 5
H5 Stronger value recognition is associated with lower perceived control, lower platform trust, and stronger support for regulation.


#Descriptive statistics
```{r descriptives}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  psych::describe() |>
  round (2)
  
```

#Normality check
```{r normality}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  lapply(shapiro.test)

```

#Knowledge-tems check
```{r knowledge-items}

dat |>
  select(s1q1_recoded:s1q6_recoded) |>
  colMeans() |>
  round(2)

```

#Reliability
```{r reliability}

kr20      <- psych::alpha(dat |> 
                          select(s1q1_recoded:s1q6_recoded), 
                          check.keys = FALSE)

beh_alpha <- psych::alpha(dat |> 
                          select(s2q1:s2q5), 
                          check.keys = FALSE)

reg_alpha <- psych::alpha(dat |> 
                          select(s4q1:s4q3), 
                          check.keys = FALSE)

r_trust  <- cor(dat$perceived_control, dat$platform_transparency, use = "complete.obs")
sb_trust <- (2 * r_trust) / (1 + r_trust)

tibble::tibble(
  Scale       = c("Awareness (6)", "Behaviour (5)", "Regulation (3)", "Trust (2)"),
  Coefficient = c("KR-20", "Cronbach alpha", "Cronbach alpha", "Spearman-Brown"),
  Value       = round(c(kr20$total$raw_alpha, beh_alpha$total$raw_alpha,
                        reg_alpha$total$raw_alpha, sb_trust), 2)
  )

```

#Correlations
```{r correlations}

corr_vars <- dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation)

sp <- psych::corr.test(corr_vars, method = "spearman")

round(sp$r, 2)   
round(sp$p, 3)   

```

#Group Comparisons
```{r group-comparisons}

outcomes <- c("awareness", "behaviour", "trust", "regulation", "value_recognition")

sapply(outcomes, function(v) wilcox.test(dat[[v]] ~ dat$gender_f)$p.value) |> round(3)

sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$age_band)$p.value)    |> round(3)
sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$edu_band)$p.value)    |> round(3)
sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$screen_band)$p.value) |> round(3)

```

```{r dunn-age}

dat |> rstatix::dunn_test(awareness ~ age_band)  |> filter(p.adj < 0.05)
dat |> rstatix::dunn_test(behaviour ~ age_band)  |> filter(p.adj < 0.05)
dat |> rstatix::dunn_test(regulation ~ age_band) |> filter(p.adj < 0.05)

print(dat |> rstatix::dunn_test(awareness ~ age_band)  |> filter(p.adj < 0.05))
print(dat |> rstatix::dunn_test(behaviour ~ age_band)  |> filter(p.adj < 0.05))
print(dat |> rstatix::dunn_test(regulation ~ age_band) |> filter(p.adj < 0.05))

```

#Regressions
```{r regression}

f <- ~ awareness + age + gender + education + screen_time

m_behaviour  <- lm(update(f, behaviour ~ .),         data = dat)
m_trust      <- lm(update(f, trust ~ .),             data = dat)
m_regulation <- lm(update(f, regulation ~ .),        data = dat)
m_value      <- lm(update(f, value_recognition ~ .), data = dat)

lapply(list(behaviour = m_behaviour, trust = m_trust,
            regulation = m_regulation, value = m_value),
       function(m) list(
         betas = effectsize::standardize_parameters(m),
         fit   = broom::glance(m)[c("r.squared","adj.r.squared","statistic","p.value")],
         vif   = car::vif(m)
       ))

```

```{r model-fit}

models <- list(behaviour = m_behaviour, trust = m_trust,
               regulation = m_regulation, value = m_value)

purrr::map_dfr(models, broom::glance, .id = "model") |>
  select(model, r.squared, adj.r.squared, statistic, p.value) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

```

```{r control-model}

m_control <- lm(perceived_control ~ value_recognition + awareness +
                  age + gender + education + screen_time, data = dat)

effectsize::standardize_parameters(m_control)
broom::glance(m_control)[c("r.squared","adj.r.squared","statistic","p.value")]

```

#Diagnostic
```{r diagnostics}

lapply(list(m_behaviour, m_trust, m_regulation, m_value),
       function(m) shapiro.test(residuals(m))$p.value)

lapply(list(m_behaviour, m_trust, m_regulation, m_value), lmtest::bptest)

which(cooks.distance(m_behaviour) > 4 / nrow(dat))

```

```{r influence-check}

infl <- c(5, 40, 52, 53, 65, 79, 81, 120)
m_beh_no_infl <- lm(update(f, behaviour ~ .), data = dat[-infl, ])
effectsize::standardize_parameters(m_beh_no_infl)

```

#Robustness
```{r robustness}

lmtest::coeftest(m_value, vcov = sandwich::vcovHC(m_value, type = "HC3"))

reg_ord <- MASS::polr(factor(s4q1) ~ awareness + age + gender + education + screen_time,
                      data = dat, Hess = TRUE)
summary(reg_ord)

```
