---
title: "Data Analysis"
author: "AG35441 - Kamila Kwiatkowska (k25120717)"
output: 
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
  )

options (scipen = 999)

library(tidyverse)
library(here)        
library(janitor)     
library(psych)       
library(rstatix)     
library(effectsize)  
library(car)         
library(lmtest)      
library(broom)
library(sandwich)  
library(ordinal)   
library(brant) 

```

```{r read-in-data}

survey_data <- read.csv("data/Analysis.csv")

dat <- survey_data |> 
  
clean_names() |>
  rename(
    awareness             = total_awareness_score,
    behaviour             = average_behaviour,
    perceived_control     = s3q1,
    value_recognition     = s3q2,
    platform_transparency = s3q3,
    trust                 = trust_score,
    regulation            = average_regulation_score,
    age                   = age_recoded,
    gender                = gender_recoded,
    education             = education_recoded,
    screen_time           = screen_time_recoded
  ) |>

mutate(
    age_band    = factor(age, levels = 1:5,
                         labels = c("18-24", "25-34", "35-44", "45-54", "55+"),
                         ordered = TRUE),
    
    edu_band    = factor(education, levels = 1:3,
                         labels = c("Secondary", "Undergraduate", "Postgraduate"),
                         ordered = TRUE),
    
    screen_band = factor(screen_time, levels = 1:4,
                         labels = c("<1h", "1-3h", "4-5h", ">5h"),
                         ordered = TRUE),
    
    gender_f    = factor(gender, levels = c(0, 1), 
                         labels = c("Male", "Female"))
    
  ) 

glimpse(dat)
cat("\nN =", nrow(dat), "\n")
    
```

```{r read-in-check}
dat |> tabyl(age_band)
dat |> tabyl(edu_band)
dat |> tabyl(screen_band)
dat |> tabyl(gender_f)
```
#Survey
This is a test analysis of a dataset on Everyday Users' Perceptions of Data Use and Value Creation in the Hidden Value Chains of the Data Economy.

#Hypotheses

##Hypothesis 1
H1 Greater awareness of data extraction is not associated with greater privacy-protective behaviour.

##Hypothesis 2
H2 Greater awareness of data extraction is associated with lower platform trust.

##Hypothesis 3
H3 Greater awareness of data extraction is associated with stronger support for regulation.

##Hypothesis 4
H4 Users recognise their everyday online activity as a source of economic value for platforms.

##Hypothesis 5
H5 Stronger value recognition is associated with lower perceived control, lower platform trust, and stronger support for regulation.


#Descriptive statistics
```{r descriptives}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  psych::describe() |>
  round (2)
  
```

#Normality check
```{r normality}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  lapply(shapiro.test)

```

#Knowledge-tems check
```{r knowledge-items}

dat |>
  select(s1q1_recoded:s1q6_recoded) |>
  colMeans() |>
  round(2)

```

#Reliability
```{r reliability}

kr20      <- psych::alpha(dat |> 
                          select(s1q1_recoded:s1q6_recoded), 
                          check.keys = FALSE)

beh_alpha <- psych::alpha(dat |> 
                          select(s2q1:s2q5), 
                          check.keys = FALSE)

reg_alpha <- psych::alpha(dat |> 
                          select(s4q1:s4q3), 
                          check.keys = FALSE)

r_trust  <- cor(dat$perceived_control, dat$platform_transparency, use = "complete.obs")
sb_trust <- (2 * r_trust) / (1 + r_trust)

tibble::tibble(
  Scale       = c("Awareness (6)", "Behaviour (5)", "Regulation (3)", "Trust (2)"),
  Coefficient = c("KR-20", "Cronbach alpha", "Cronbach alpha", "Spearman-Brown"),
  Value       = round(c(kr20$total$raw_alpha, beh_alpha$total$raw_alpha,
                        reg_alpha$total$raw_alpha, sb_trust), 2)
  )

```

#Correlations
```{r correlations}

corr_vars <- dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation)

sp <- psych::corr.test(corr_vars, method = "spearman")

round(sp$r, 2)   
round(sp$p, 3)   

```

#Group Comparisons
```{r group-comparisons}

outcomes <- c("awareness", "behaviour", "trust", "regulation", "value_recognition")

sapply(outcomes, function(v) wilcox.test(dat[[v]] ~ dat$gender_f)$p.value) |> round(3)

sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$age_band)$p.value)    |> round(3)
sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$edu_band)$p.value)    |> round(3)
sapply(outcomes, function(v) kruskal.test(dat[[v]] ~ dat$screen_band)$p.value) |> round(3)

```

```{r dunn-age}

dat |> rstatix::dunn_test(awareness ~ age_band)  |> filter(p.adj < 0.05)
dat |> rstatix::dunn_test(behaviour ~ age_band)  |> filter(p.adj < 0.05)
dat |> rstatix::dunn_test(regulation ~ age_band) |> filter(p.adj < 0.05)

print(dat |> rstatix::dunn_test(awareness ~ age_band)  |> filter(p.adj < 0.05))
print(dat |> rstatix::dunn_test(behaviour ~ age_band)  |> filter(p.adj < 0.05))
print(dat |> rstatix::dunn_test(regulation ~ age_band) |> filter(p.adj < 0.05))

```

#Regressions
```{r regression}

f <- ~ awareness + age + gender + education + screen_time

m_behaviour  <- lm(update(f, behaviour ~ .),         data = dat)
m_trust      <- lm(update(f, trust ~ .),             data = dat)
m_regulation <- lm(update(f, regulation ~ .),        data = dat)
m_value      <- lm(update(f, value_recognition ~ .), data = dat)

lapply(list(behaviour = m_behaviour, trust = m_trust,
            regulation = m_regulation, value = m_value),
       function(m) list(
         betas = effectsize::standardize_parameters(m),
         fit   = broom::glance(m)[c("r.squared","adj.r.squared","statistic","p.value")],
         vif   = car::vif(m)
       ))

```

```{r model-fit}

models <- list(behaviour = m_behaviour, trust = m_trust,
               regulation = m_regulation, value = m_value)

purrr::map_dfr(models, broom::glance, .id = "model") |>
  select(model, r.squared, adj.r.squared, statistic, p.value) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

```

```{r control-model}

m_control <- lm(perceived_control ~ value_recognition + awareness +
                  age + gender + education + screen_time, data = dat)

effectsize::standardize_parameters(m_control)
broom::glance(m_control)[c("r.squared","adj.r.squared","statistic","p.value")]

```

#Diagnostic
```{r diagnostics}

lapply(list(m_behaviour, m_trust, m_regulation, m_value),
       function(m) shapiro.test(residuals(m))$p.value)

lapply(list(m_behaviour, m_trust, m_regulation, m_value), lmtest::bptest)

which(cooks.distance(m_behaviour) > 4 / nrow(dat))

```

```{r influence-check}

infl <- c(5, 40, 52, 53, 65, 79, 81, 120)
m_beh_no_infl <- lm(update(f, behaviour ~ .), data = dat[-infl, ])
effectsize::standardize_parameters(m_beh_no_infl)

```

#Robustness
```{r robustness}

lmtest::coeftest(m_value, vcov = sandwich::vcovHC(m_value, type = "HC3"))

reg_ord <- MASS::polr(factor(s4q1) ~ awareness + age + gender + education + screen_time,
                      data = dat, Hess = TRUE)
summary(reg_ord)

```

# Visualizations

```{r plot-theme}

library(scales)

theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 14),
      plot.subtitle = element_text(hjust = 0.5),
      plot.margin   = margin(10, 15, 10, 10),
      strip.text    = element_text(face = "bold")
    )
)

fill_main <- "#2C6E8F"   
accent    <- "#B4460E"   

```

## Knowledge item difficulty

```{r knowledge-difficulty}

tibble::tibble(
  item = factor(
    c("Q1 Cookie consent", "Q2 App deletion", "Q3 CAPTCHA/AI",
      "Q4 Reject breaks site", "Q5 Third-party sharing", "Q6 Incognito"),
    levels = rev(c("Q1 Cookie consent", "Q2 App deletion", "Q3 CAPTCHA/AI",
                   "Q4 Reject breaks site", "Q5 Third-party sharing", "Q6 Incognito"))
  ),
  
  correct = colMeans(dat |> select(s1q1_recoded:s1q6_recoded))
) |>
  
  ggplot(aes(item, correct)) +
  geom_col(fill = fill_main) +
  geom_text(aes(label = scales::percent(correct, accuracy = 1)), hjust = -0.15, size = 3.5) +
  coord_flip(ylim = c(0, 1)) +
  labs(x = NULL, 
       y = "Proportion correct", 
       title = "Awareness by knowledge item")

```

## Composite distributions

```{r fig-distributions}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  pivot_longer(everything(), names_to = "measure", values_to = "score") |>
  ggplot(aes(score)) +
  geom_histogram(bins = 12, fill = fill_main) +
  facet_wrap(~ measure, scales = "free") +
  labs(x = "Score", 
       y = "Count", 
       title = "Distribution of composite measures")

```

## Value recognition

```{r fig-value}

dat |>
  count(value_recognition) |>
  
  ggplot(aes(factor(value_recognition), n, fill = factor(value_recognition))) +
  
  geom_col() +
  
  scale_fill_brewer(palette = "Blues") +
  
  labs(x = "Value recognition (1 = not at all, 5 = a great deal)",
       y = "Respondents", 
       title = "Recognition of value generation") +
  
  theme(legend.position = "none")

```

## Scatterplot: Awareness and behaviour

```{r fig-h1}

ggplot(dat, aes(awareness, behaviour)) +
  
  geom_jitter(width = 0.15, height = 0.05, alpha = 0.5, colour = "#4C4C4C") +
  
  geom_smooth(method = "lm", 
              se = TRUE, 
              colour = accent, 
              fill = accent, 
              alpha = 0.15) +
  
  labs(x = "Awareness score (0-6)", 
       y = "Protective behaviour (1-5)",
       title = "Awareness and protective behaviour")

```

## Scatterplot: Value recognition and perceived control

```{r fig-h5-control}

ggplot(dat, aes(value_recognition, perceived_control)) +
  
  geom_jitter(width = 0.1, 
              height = 0.1, 
              alpha = 0.5, 
              colour = "#4C4C4C") +
  
  geom_smooth(method = "lm", 
              se = TRUE, 
              colour = accent, 
              fill = accent, 
              alpha = 0.15) +
  
  labs(x = "Value recognition (1-5)", 
       y = "Perceived control (1-5)",
       title = "Value recognition and perceived control")

```

## Scatterplot: Value recognition and regulation

```{r fig-h5-reg}

ggplot(dat, aes(value_recognition, regulation)) +
  
  geom_jitter(width = 0.1, 
              height = 0.05, 
              alpha = 0.5, 
              colour = "#4C4C4C") +
  
  geom_smooth(method = "lm", 
              se = TRUE, 
              colour = accent, 
              fill = accent, 
              alpha = 0.15) +
  
  labs(x = "Value recognition (1-5)", 
       y = "Support for regulation (1-5)",
       title = "Value recognition and support for regulation")

```

## Scatterplot: Behaviour and regulation

```{r fig-beh-reg}

ggplot(dat, aes(behaviour, regulation)) +
  
  geom_jitter(width = 0.05, 
              height = 0.05, 
              alpha = 0.5, 
              colour = "#4C4C4C") +
  
  geom_smooth(method = "lm", 
              se = TRUE, 
              colour = accent, 
              fill = accent, 
              alpha = 0.15) +
  
  labs(x = "Protective behaviour (1-5)", 
       y = "Support for regulation (1-5)",
       title = "Protective behaviour and support for regulation")

```

## Boxplot: Age effects

```{r fig-age}

dat |>
  select(age_band, awareness, behaviour, regulation) |>
  
  pivot_longer(-age_band, names_to = "measure", values_to = "score") |>
  
  ggplot(aes(age_band, score, fill = age_band)) +
  
  geom_boxplot(alpha = 0.85) +
  
  facet_wrap(~ measure, scales = "free_y") +
  
  scale_fill_brewer(palette = "Blues") +
  
  labs(x = "Age", 
       y = "Score", 
       title = "Outcomes by age band") +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

```

## Boxplot: Regulation by education

```{r fig-edu-reg}

ggplot(dat, aes(edu_band, regulation, fill = edu_band)) +
  
  geom_boxplot(alpha = 0.85) +
  
  scale_fill_brewer(palette = "Greens") +
  
  labs(x = "Education", 
       y = "Support for regulation (1-5)",
       title = "Support for regulation by education") +
  
  theme(legend.position = "none")

```

## Correlation heatmap

```{r fig-heatmap}

m <- dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  
  cor(method = "spearman")

as.data.frame(as.table(m)) |>
  ggplot(aes(Var1, Var2, fill = Freq)) +
  
  geom_tile(colour = "white") +
  
  geom_text(aes(label = round(Freq, 2)), size = 3) +
  
  scale_fill_gradient2(low = "#B2182B", mid = "white", high = "#2166AC",
                       limits = c(-1, 1), name = "rho") +
  labs(x = NULL, 
       y = NULL, 
       title = "Spearman correlations") +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Coefficient forest plot

```{r fig-coef}

models <- list(Behaviour = m_behaviour, Trust = m_trust,
               Regulation = m_regulation, Value = m_value)

purrr::map_dfr(models, function(m) {
  sp <- effectsize::standardize_parameters(m)
  tibble::tibble(term = sp$Parameter, beta = sp$Std_Coefficient,
                 low = sp$CI_low, high = sp$CI_high)
  
}, .id = "model") |>
  filter(term != "(Intercept)") |>
  
  ggplot(aes(beta, term, colour = model)) +
  
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  
  geom_pointrange(aes(xmin = low, xmax = high)) +
  
  facet_wrap(~ model) +
  
  scale_colour_brewer(palette = "Set2") +
  
  labs(x = "Standardised coefficient (95% CI)", 
       y = NULL,
       title = "Predictors across the four models") +
  
  theme(legend.position = "none")

```

## Awareness effect across outcomes

```{r fig-awareness-effect}

models <- list(Behaviour = m_behaviour, Trust = m_trust,
               Regulation = m_regulation, Value = m_value)

purrr::map_dfr(models, function(m) {
  sp <- effectsize::standardize_parameters(m)
  tibble::tibble(term = sp$Parameter, beta = sp$Std_Coefficient,
                 low = sp$CI_low, high = sp$CI_high)
  
}, .id = "model") |>
  
  filter(term == "awareness") |>
  
  ggplot(aes(beta, model, colour = model)) +
  
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  
  geom_pointrange(aes(xmin = low, xmax = high)) +
  
  scale_colour_brewer(palette = "Set2") +
  
  labs(x = "Standardised awareness coefficient (95% CI)", 
       y = NULL,
       title = "Effect of awareness on each outcome") +
  
  theme(legend.position = "none")

```