---
title: "Data Analysis"
author: "AG35441 - Kamila Kwiatkowska (k25120717)"
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
  )

options (scipen = 999)

library(tidyverse)
library(here)        
library(janitor)     
library(psych)       
library(rstatix)     
library(effectsize)  
library(car)         
library(lmtest)      
library(broom)
library(sandwich)  
library(ordinal)   
library(brant) 

```

```{r read-in-data}

survey_data <- read.csv("data/Analysis.csv")

dat <- survey_data |> 
  
clean_names() |>
  rename(
    awareness             = total_awareness_score,
    behaviour             = average_behaviour,
    perceived_control     = s3q1,
    value_recognition     = s3q2,
    platform_transparency = s3q3,
    trust                 = trust_score,
    regulation            = average_regulation_score,
    age                   = age_recoded,
    gender                = gender_recoded,
    education             = education_recoded,
    screen_time           = screen_time_recoded
  ) |>

mutate(
    age_band    = factor(age, levels = 1:5,
                         labels = c("18-24", "25-34", "35-44", "45-54", "55+"),
                         ordered = TRUE),
    
    edu_band    = factor(education, levels = 1:3,
                         labels = c("Secondary", "Undergraduate", "Postgraduate"),
                         ordered = TRUE),
    
    screen_band = factor(screen_time, levels = 1:4,
                         labels = c("<1h", "1-3h", "4-5h", ">5h"),
                         ordered = TRUE),
    
    gender_f    = factor(gender, levels = c(0, 1), 
                         labels = c("Male", "Female"))
    
  ) 

glimpse(dat)
cat("\nN =", nrow(dat), "\n")
    
```

```{r read-in-check}
dat |> tabyl(age_band)
dat |> tabyl(edu_band)
dat |> tabyl(screen_band)
dat |> tabyl(gender_f)
```
#Survey
This is a test analysis of a dataset on 

#Hypotheses

##Hypothesis 1
H1 Greater awareness of data extraction is not associated with greater privacy-protective behaviour.

##Hypothesis 2
H2 Greater awareness of data extraction is associated with lower platform trust.

##Hypothesis 3
H3 Greater awareness of data extraction is associated with stronger support for regulation.

##Hypothesis 4
H4 Users recognise their everyday online activity as a source of economic value for platforms.

##Hypothesis 5
H5 Stronger value recognition is associated with lower perceived control, lower platform trust, and stronger support for regulation.


#Descriptive statistics
```{r descriptives}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  psych::describe() |>
  round (2)
  
```

```{r normality}

dat |>
  select(awareness, behaviour, perceived_control, value_recognition,
         platform_transparency, trust, regulation) |>
  lapply(shapiro.test)

```

```{r knowledge-items}

dat |>
  select(s1q1_recoded:s1q6_recoded) |>
  colMeans() |>
  round(2)

```
